---
title: "HW04 Part 2: Complete Chapter 3"
author: "Donovan Kleinberg"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
```

Data for 1.

```{r}
data(trees)
?trees

### Plot 1
ggplot(data = trees, mapping = aes(x = 	Height, y = Girth)) +
geom_point()
```

Comments for Plot 1: It appears that there is a slight correlation that as height increases girth increases as well.

Data for 2.

```{r}
data(lynx)
?lynx

### Plot 2
year <- (1821:1934)
pelts <- as.vector(lynx)
lynx_pelts <- data.frame(year, pelts)
ggplot(data = lynx_pelts, mapping = aes(x = year, y = pelts)) +
  geom_line(color = "maroon")
```

Comments for Plot 2: There are peaks every 5 years or so with massive peaks every 20 years. If I remember correctly early explanations offered for these peaks in lynx were crazy things like sunspot cycles and the like.

Data for 3.

```{r}
data(iris)
?iris

### Plot 3
ggplot(data = iris, mapping = aes(x = Petal.Length, y = Petal.Width)) +
  geom_point(mapping = aes(color = "Species", shape = "Species", size = 2)) +
  xlab("Petal Length") + ylab("Petal Width")
```

Comments for plot 3: Overall there seems to  be a very strong correlation between increasing petal length and increasing petal width. There is also an interesting gap at around the 3 mark for petal length, perhaps a peculiarity of iris plants?

Data for 4.