---
title: "HW04 Part 2: Complete Chapter 3"
author: "Donovan Kleinberg"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
```

Data for 1.

```{r}
data(trees)
?trees

### Plot 1
ggplot(data = trees, mapping = aes(x = 	Height, y = Girth)) +
geom_point()
```

Comments for Plot 1: It appears that there is a slight correlation that as height increases girth increases as well.

Data for 2.

```{r}
data(lynx)
?lynx

### Plot 2
year <- (1821:1934)
pelts <- as.vector(lynx)
lynx_pelts <- data.frame(year, pelts)
ggplot(data = lynx_pelts, mapping = aes(x = year, y = pelts)) +
  geom_line(color = "maroon")
```

Comments for Plot 2: There are peaks every 5 years or so with massive peaks every 20 years. If I remember correctly early explanations offered for these peaks in lynx were crazy things like sunspot cycles and the like.

Data for 3.

```{r}
data(iris)
?iris

### Plot 3
ggplot(data = iris, mapping = aes(x = Petal.Length, y = Petal.Width)) +
  geom_point(mapping = aes(color = "Species", shape = "Species", size = 2)) +
  xlab("Petal Length") + ylab("Petal Width")
```

Comments for plot 3: Overall there seems to  be a very strong correlation between increasing petal length and increasing petal width. There is also an interesting gap at around the 3 mark for petal length, perhaps a peculiarity of iris plants?

Data for 4 and 5.

```{r}
msleep
?msleep
```

Plots

```{r}
### Plot 4
ggplot(data = msleep) +
  geom_violin(mapping = aes(x = vore, y =sleep_rem), fill = "gray30")

### Plot 5
ggplot(data = msleep) +
  geom_violin(mapping = aes(x = vore, y =sleep_rem), fill = "gray70")
```

Comments for plots 4 and 5: Violin plots show the distribution of data as a shape of varying thickness with thicker portions being where more data is located at. Gray 30 was much darker than Gray 70.

Data for plot 7.

```{r}
### Plot 7
ggplot(data = msleep, mapping = aes(x = order, y = sleep_total)) +
  geom_boxplot() +
  coord_flip()

```

Comments for plot 7: It appears that some orders of species have much higher variation in the amount of sleep they have compared to other species which have practically no variation.

Data for plot 8.

```{r}
ggplot(data = msleep, mapping = aes(x = conservation, y = awake)) +
  geom_boxplot(mapping = aes(color = conservation)) +
  geom_point(mapping = aes(color = conservation)) +
  geom_jitter(mapping = aes(color = conservation)) +
  xlab("Conservation") + ylab("Awake") +
  scale_color_discrete(name = "Conservation")
```

Comments for plot 8: It seems like there is great variation among conservation types in their periods of being awake.

Data for plots 9 and 10.

```{r}
ggplot(data = msleep, mapping = aes(x =sleep_cycle, y = bodywt)) +
  geom_point()
```




