---
title: "HW05 Part 2: Importing"
author: "Donovan Kleinberg"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Limpets 

```{r}
library(tidyverse)
file_path <- "C:\\Misc. Files\\BI485 Folder\\Kleinberg_Donovan\\data"
limpet <- read_csv(file.path(file_path, "limpets.csv"))
```
```{r}
limpet
```

Chunk for column identification.

```{r}
read_csv(file.path(file_path, "limpets.csv"), col_types = c("ncc"))
```

Chunk for ordering site

```{r}
Ordered_sites <- factor(limpet, levels = c("PBL", "KNRM", "VBG", "WP", "PF", "DP", "SIO", "CTZ1", "CTZ2", "CNM"), ordered = TRUE)
```

Chunk for boxplot

```{r}
ggplot(data = limpet, mapping = aes(x = Site, y = Length)) +
  geom_boxplot(mapping = aes(color = Protected)) +
    xlab("Site") + ylab("Length (mm)")
```

Description for Boxplot: There are 2 sites with outliers, DP and WP. The protected sites all seem to have longer limpets than non protected sites.

## Terns

```{r}
file_path <- "C:\\Misc. Files\\BI485 Folder\\Kleinberg_Donovan\\data"
tern <- read.delim(file.path(file_path, "roseate_terns.txt"))
tern
```

Tidying the data

```{r}
tidy_tern <- tern %>% 
  gather("AR", "AL", "AC", "BR", "BL", "BC", "JR", "JL", "JC", key = "Site", value = "Individuals")
tidy_tern
```

Filtering out NA data

```{r}
filter(tidy_tern, is.na(Individuals) == FALSE)
```

Lineplot of Tern data

```{r}
ggplot(data = tidy_tern, mapping = aes(x = year, y = Individuals)) +
  geom_line(mapping = aes(color = Individuals)) +
  xlab("Year") + ylab("Number of Tern Per Site") +
    facet_wrap(~ Site, nrow = 3)
```

Description for line plot: The site labelled AR has shown an obvious increase in population size, sites BR and JR meanwhile have shown obvious decreases in population size. Gaps in the data symbolize NA data due to no survey during that period.

## Darters

```{r}
file_path <- "C:\\Misc. Files\\BI485 Folder\\Kleinberg_Donovan\\data"
darters <- read.delim(file.path(file_path, "darter.txt"))
darters

Column_start <- c(1, 4, 15, 16, 17, 18, 21, 25, 29, 33, 36, 39, 40)
Column_stop <- c(3, 14, 15, 16, 17, 20, 24, 28, 32, 35, 38, 40, 41)
names <- c("specimen id", "species", "riffle", "age", "sex", "length", "depth", "velocity", "temperature", "majsub", "minsub", "majtype", "mintype")
tidy_darters <- read_fwf(file.path(file_path, "darter.txt"), fwf_positions(Column_start, Column_stop, col_names = names), skip = 15)
tidy_darters
```

Unordered and Ordered Factors

```{r}
riff <- c("1", "2")
riffle <- factor(riff)
major <- c("s", "fg", "sg", "lg", "c")
major_type <- factor(major, ordered = TRUE, levels = c("s", "fg", "sg", "lg", "c"))
```

The Pipe

```{r}
good_darters <- tidy_darters %>%
  filter(!species %in% c("zonale", "tetrazonum"))
```

```{r}
gooder_darters <- subset(good_darters, select = -c(mintype,minsub) )
gooder_darters

colnames(gooder_darters)[colnames(gooder_darters)=="majsub"] <- "major_substrate"
colnames(gooder_darters)[colnames(gooder_darters)=="majtype"] <- "major_type"
gooder_darters

gooder_darters %>% arrange(-desc(`specimen id`))
```

Plot 1

```{r}
ggplot(data = gooder_darters, mapping = aes(x = length, y = depth)) + 
  geom_point(mapping = aes(color = species, shape = species)) +
  facet_wrap(~ riffle)
```

Description of plot 1: Riffle one has a higher amount of spectabile and also has fish being found in more shallow water while riffle 2 has more blenniodies, larger fish and fish that stay deeper.

Plot 2

```{r}
ggplot(data = gooder_darters, mapping = aes(x = species, y = length)) + 
  geom_boxplot() +
  facet_wrap(~ riffle)

```

Description of plot 2: Riffle 2 has by far the greater number of outliers. 

Plot 3

```{r}
ggplot(data = gooder_darters, mapping = aes(x = species, y = length)) + 
  geom_boxplot() +
  facet_grid(species ~ riffle)

ggplot(data = gooder_darters, mapping = aes(x = species, y = length)) + 
  geom_boxplot() +
  facet_grid(riffle ~ species)
```

Description of plot 3: Switching species and riffle results in a graph that is less pleasing to read and one that doesn't appear to function properly.

## Abalone

Chunk 1

```{r}
file_path <- "C:\\Misc. Files\\BI485 Folder\\Kleinberg_Donovan\\data"
abalone <-  read_csv(file.path(file_path, "abalone.csv"))
abalone
abalone_no_first <- abalone[-c(1)]
abalone_no_first

ggplot(data = abalone_no_first, mapping = aes(x = Type, y = Height)) +
  geom_boxplot()
```

Chunk 2

```{r}
ggplot(data = abalone_no_first, mapping = aes(x = Height, y = Rings)) +
  geom_point()
```

Description for chunk 2: It would appear that the outliers are only around 7-10 years old and are not extremely old.

Chunk 3

```{r}
abalone_error <- filter(abalone_no_first, Height < 0.3)
abalone_ordered <- abalone_error$Type <- factor(abalone_error$Type, levels = c("I", "F", "M"), ordered = TRUE)
abalone_ordered

ggplot(data = abalone_error, mapping = aes(x = Rings, y = Height)) +
  geom_point() +
  facet_wrap("Type")
```

Description for chunk 3: Overall males and females are larger than immature abalone for the same number of rings. Overall the females tend to be slightly bigger with neither gender truely becoming totally seperate in size from immature specimens.

Chunk 4

```{r}
filter(abalone_error, Height == 0)
abalone_error[abalone_error == 0] <- NA

ggplot(data = abalone_error, mapping = aes(x = Rings, y = Height)) +
  geom_point() +
  facet_wrap("Type")
```

Chunk 5

```{r}
ggplot(data = abalone_error, mapping = aes(x = Diameter, y = WholeWeight)) +
  geom_point(mapping = aes(color = Type)) +
  ylab("Whole Weight")

ggplot(data = abalone_error, mapping = aes(x = LongestShell, y = Rings)) +
  geom_point(mapping = aes(color = Type, shape = Type)) +
  xlab("Longest Shell")
```

