---
title: "HW 08 Part 1"
author: "Donovan Kleinberg"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(lubridate)
```

Part 1.

This chunk has the approved copy paste

```{r}
birth_date <- "2/10/1997"

# Elizabeth Garret Anderson was the first female physician
anderson <- "9 June 1836"

# Mary Anning discovered the first complete Plesiosaur fossil, 
# and other important Jurassic fossils
anning <-  "1799/05/21"

# Alice Augusta Ball was a chemist who developed the
# first injectable treatment for leprosy.
ball <- "July 24, 1892"

# Sylvia Earle was the first chieft scientist at NOAA
earle <- "August, 30 (1935)"

# Rosalind Franklin used x-ray diffraction to 
# discover DNA structure. Watson and Crick claimed credit.
franklin <- "25th of July, 1920"

# Ester Lederberg was the first to discover that viruses 
# infect bacteria. Led to discovery of lambda bacteriophage
lederberg <- "December 18, 1922"

# Barbara McClintock discovered transposable elements (jumping genes),
# and developed staining techniques to study chromosomes
mcclintock <- "16061902"

# Martin Luther King's birthday is coming up. Let's remember a dream
mlk_birth <- "1/15/29"
mlk_dream <- "1963 August 08"
```

Answer for 1.

```{r}
birthday <- mdy(birth_date)
anderson <- dmy(anderson)
anning <- ymd(anning)
ball <- mdy(ball)
earle <- mdy(earle)
franklin <- dmy(franklin)
lederberg <- mdy(lederberg)
mcclintock <- dmy(mcclintock)
mlk_birth <- mdy(mlk_birth)
mlk_dream <- ymd(mlk_dream)

birthday
anderson
anning
ball
earle
franklin
lederberg
mcclintock
mlk_birth
mlk_dream
```

Answer for 2.

```{r}
mlk_birth <- update(mlk_birth, year = 1929)

mlk_birth
```

Answer for 3.

```{r}
wday(birthday, label = TRUE, abbr = FALSE)

# I was apparently born on a Monday, figured as much. 
```

Answer for 4.

```{r}
present <- now()

year <- year(present)

present_birthday <- update(birthday, year = year)

present_birthday

wday(present_birthday, label = TRUE, abbr = FALSE)
```

Answer for 5.

```{r}
future_birthday <- update(birthday, year = 2525)
wday(future_birthday, label = TRUE, abbr = TRUE)

justforfun <- future_birthday - birthday

justforfun

# That's a lot of days.

future_age <- year(future_birthday) - year(birthday)

future_age
```

Answer for 6.

```{r}
time_dif <- ball - anderson

time_dif
```

Answer for 7.

```{r}
mlk_time_dif <- mlk_dream - mlk_birth
mlk_dur <- as.duration(mlk_time_dif)

mlk_dur
```

Answer for 8.

```{r}
birth_dates_vector <- c(birthday, anderson, anning, ball, earle, franklin, lederberg, mcclintock, mlk_birth)
names(birth_dates_vector) <- c("Kleinberg", "Anderson", "Anning", "Ball", "Earle", "Franklin", "Lederberg", "McClintock", "King")

birth_dates_vector

## Leap Years 

leap_year <- leap_year(birth_dates_vector)
year(birth_dates_vector[leap_year])

## Time Between My Birth and The Others

days_passed <- time_length(birth_dates_vector - birthday, "day")
sort(days_passed, decreasing = FALSE)
```

Part 2.

Lake Ice Data.

```{r}
file_path <- "C:\\Misc. Files\\BI485 Folder\\Kleinberg_Donovan\\data"
lake_unfiltered <- read_csv(file.path(file_path, "lake_ice.csv"))

lake_unfiltered
```

Chunk filters Mendota

```{r}
lake_ice <- filter(lake_unfiltered, lakename == "LAKE MENDOTA")

lake_ice
```

Chunk selects specified columns.

```{r}
lake_ice_sel <- select(lake_ice, iceon_year, iceon_month, iceon_day, iceoff_year, iceoff_month, iceoff_day)

lake_ice_sel
```

Chunk gathers data into 2 columns.

```{r}
lake_ice_grouped <- lake_ice_sel %>%
  select(iceon_year, iceon_month, iceon_day, iceoff_year, iceoff_month, iceoff_day)%>%
  mutate(ice_on = make_date(iceon_year, iceon_month, iceon_day),
         ice_off = make_date(iceoff_year, iceoff_month, iceoff_day))

lake_ice_grouped
```

Chunk creates duration

```{r}
lake_ice_duration <- lake_ice_grouped %>%
  mutate(duration = as.numeric(ice_off - ice_on))

lake_ice_duration

### It would appear the duration is listed in days.
```

Chunk has line plot

```{r}
ggplot(lake_ice_duration, aes(x = iceoff_year, y = duration), na.rm = TRUE) +
  geom_line() +
  coord_cartesian(xlim = c(1853, 2013)) +
  labs(x = "Year", y = "Ice Cover Duration per Year (Days)")
```

Soil Data

```{r}
file_path <- "C:\\Misc. Files\\BI485 Folder\\Kleinberg_Donovan\\data"
soil_raw <- read_tsv(file.path(file_path, "soil_co2.txt"))

soil_raw
```

Chunk has date variable

```{r}
date_var <- make_date(year = 2006, day = 1, month = 1)
```

Chunk has renamed columns

```{r}
soil_renamed <- soil_raw %>%
  rename(Day_number = `DayN[]`,
         Day_sampled = `Datecode[]`,
         Minutes = `Min[]`,
         Vegetation = `Veg[]`,
         Plot = `Plot[]`,
         Group = `Group[]`,
         Corrected_CO2_Level = `CorrCO2[]`,
         Soil_temp = `SoilT[]`)

### I'm assuming the soil temp is measured in kelvin which fits well, otherwise I don't know what it is. 

soil_renamed
```

Chunk adds data and groups by year and month

```{r}
date_var_additional <- ymd(date_var)

soil_date_added <- soil_renamed %>%
  mutate(Date = date_var_additional + Day_number, Year = year(Date), Month = month(Date, label = TRUE, abbr = TRUE)) %>%
  group_by(Year, Month)

soil_date_added
```

Chunk summarizes 

```{r}
soil_summ <- soil_date_added %>%
  summarise(    Mean_CO2 = mean(Corrected_CO2_Level), StdECO2 = sd(Corrected_CO2_Level)/sqrt(n()))

soil_summ
```

Chunk has column chart

```{r}
ggplot(data = soil_summ, aes(x = Month, y = Mean_CO2)) +
  geom_col() +
  geom_errorbar(aes(ymin = Mean_CO2 - StdECO2, ymax = Mean_CO2 + StdECO2, width = 0.1)) +
  facet_wrap(~Year, nrow = 2)
```

Chunk has point plot

```{r}
ggplot(data = soil_summ, aes(x = Month, y = Mean_CO2)) +
  geom_point() +
  geom_linerange(aes(ymin = Mean_CO2 - StdECO2, ymax = Mean_CO2 + StdECO2)) +
  facet_wrap(~Year, nrow = 2)
```

Mississippi River Data

```{r}
file_path <- "C:\\Misc. Files\\BI485 Folder\\Kleinberg_Donovan\\data"
thebes_discharge_raw <- read_csv(file.path(file_path, "thebes_discharge.csv"))

thebes_discharge_raw
```

