---
title: "HW 07 Part 1"
author: "Donovan Kleinberg"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
install.packages("GGally")
```

```{r}
library(GGally)
install.packages("devtools")
library(devtools)
```

```{r}
library(patchwork)
```

Loading sparrows data

```{r}
file_path <- "C:\\Misc. Files\\BI485 Folder\\Kleinberg_Donovan\\data"
sparrow_raw <- read_tsv(file.path(file_path, "sparrows.txt"), col_names = TRUE)
sparrow_raw 
```

Chunk adds rows needed.

```{r}
sparrows <- sparrow_raw %>%
  mutate(box_group = "Sparrows",
         row_order = 1:nrow(.))

View(sparrows)
```

Chunk creates initial boxplot

```{r}
plot_box <- sparrows %>%
  ggplot(aes(x = box_group, y = wingcrd)) +
  geom_boxplot(width = 0.25) +
  labs(x = NULL, y = "Wing Length (mm)")
```

Chunk adds cleveland plot

```{r}
plot_cleveland <- sparrows %>%
  ggplot(aes(x = wingcrd, y = row_order)) +
  geom_point() +
  labs(x = "Wing Length (mm)", y = "Order of the data")
```

Chunk is stitching the plots together

```{r}
plot_box + plot_cleveland + coord_flip()
```
Figure 3

Chunk begins work on figure 3 by creating all plots and assigning them to variables

```{r}
p_culmen <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = culmen)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

p_culmen

p_nalospi <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = nalospi)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

p_nalospi

p_wt <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wt)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

p_wingcrd <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wingcrd)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

p_tarsus <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = tarsus)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

p_head <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = head)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

p_culmen + p_nalospi + p_wt + p_wingcrd + p_tarsus + p_head + plot_layout(nrow = 2)
```

Answer for question regarding figure 3: It would appear the the head, tarsus and nalospi are the variables with extreme outliers.

Figure 5

Chunk plots out histogram

```{r}
h1 <- sparrows %>%
    filter(Month %in% 6:8) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by = 2))

h1
```

Chunk plots out second facet histograms

```{r}
h2 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  mutate(Month = case_when(
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August")) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by = 2)) +
  facet_wrap(~ Month, ncol = 1)

h2
```

Chunk puts h1 and h2 together

```{r}
h1 + h2
```

Figure 10

Chunk does scatmat

```{r}
sparrows %>% ggscatmat(columns = 1:7)
```

Figure 11

Chunk runs specified code

```{r}
month_order = c("May", "June", "July", "August", "September")
sparrows %>% 
  filter(Month %in% 5:9,
         Sex != 0) %>% 
  mutate(Month = case_when(
    Month == 5 ~ "May",
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August",
    Month == 9 ~ "September"),
    Month = factor(Month, levels = month_order, ordered = TRUE),
    Sex = ifelse(Sex == 4, "Male", "Female")) %>% 
  ggplot(aes(x = wingcrd,
             y = wt)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(Sex ~ Month)
```

PART 2
Crabs

```{r}
raw_crab <- read_csv(file.path(file_path, "aegla_crabs.csv"))

raw_crab
```

Chunk selects and removes columns

```{r}
crabs_selected <- select(raw_crab, -Site, -AT, -WT, -Shrimp, -Aeglam, -AeglaFe, -AeglaFo)

crabs_selected
```

Chunk removes rows with NA.

```{r}
crabsclean <- drop_na(crabs_selected)
```

Chunk adds dummy variable.

```{r}
crabs_wrang <- crabs_clean %>%
   mutate(row_order = 1:nrow(.))

crabs_wrang
```

Chunk creates variables to explore the data

```{r}
p_width <- crabs_wrang %>%
  ggplot(aes(x =Width))+
  geom_point(aes(y = row_order))

p_depth <- crabs_wrang %>% 
  ggplot(aes(x = Depth))+
  geom_point(aes(y = row_order))

p_flow <- crabs_wrang %>%
  ggplot(aes(x = Flow))+
  geom_point(aes(y = row_order))

p_ph <- crabs_wrang %>%
  ggplot(aes(x = pH))+
  geom_point(aes(y = row_order))

p_tds <- crabs_wrang %>%
  ggplot(aes(x = TDS))+
  geom_point(aes(y = row_order))

p_cond<- crabs_wrang %>%
  ggplot(aes(x = Cond))+
  geom_point(aes(y = row_order))

p_n <- crabs_wrang %>%
  ggplot(aes(x = N))+
  geom_point(aes(y = row_order))

p_ni <- crabs_wrang %>%
  ggplot(aes(x = Ni))+
  geom_point(aes(y = row_order))

p_nt <- crabs_wrang %>%
  ggplot(aes(x = Nt))+
  geom_point(aes(y = row_order))

p_phos <- crabs_wrang %>%
  ggplot(aes(x = Phos))+
  geom_point(aes(y = row_order))

p_mg <- crabs_wrang %>%
  ggplot(aes(x = Mg))+
  geom_point(aes(y = row_order))

p_ca <- crabs_wrang %>%
  ggplot(aes(x = Ca))+
  geom_point(aes(y = row_order))

p_crab <- crabs_wrang %>%
  ggplot(aes(x = Crab))+
  geom_point(aes(y = row_order))
```

Chunk contains variables plotted to determine extreme outliers. 

```{r}
p_width

p_depth

p_flow

p_ph

p_tds

p_cond

p_n

p_ni

p_nt

p_phos

p_mg

p_ca

p_crab
```

Discussion regarding outliers. I believe Depth and pH both have data that somewhat resembles outliers with depth having a single measurement 2 meters deeper than the others and pH having a measurement around row order 15 that is a little lower than any other. However, Phos has the one true outlier with a measurement near row 0 that is much greater, around 0.75 compared to the rest which hover near 0.06 - 0.12.

Chunk has the 4 Cleveland plots.

```{r}
### Cleveland 1

cleve_one <-  crabs_wrang %>%
  ggplot(aes(x = Phos, y = row_order)) +
  geom_point() +
  labs(x = "Phosphorus", y = "Order of the data")

cleve_one

### Cleveland 2

cleve_two <-  crabs_wrang %>%
  ggplot(aes(x = Width, y = row_order)) +
  geom_point() +
  labs(x = "Width (cm)", y = "Order of the data")

cleve_two

### Cleveland 3

cleve_three <-  crabs_wrang %>%
  ggplot(aes(x = N, y = row_order)) +
  geom_point() +
  labs(x = "N", y = "Order of the data")

cleve_three

### Cleveland 4

cleve_four <-  crabs_wrang %>%
  ggplot(aes(x = TDS, y = row_order)) +
  geom_point() +
  labs(x = "TDS", y = "Order of the data")

cleve_four
```

Chunk uses patchwork to stitch the 4 clevelands together

```{r}
cleve_one + cleve_two + cleve_three + cleve_four + plot_layout(ncol = 2)
```

