---
title: "HW 07 Part 1"
author: "Donovan Kleinberg"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
install.packages("GGally")
```

```{r}
library(GGally)
install.packages("devtools")
library(devtools)
```

```{r}
library(patchwork)
```

Loading sparrows data

```{r}
file_path <- "C:\\Misc. Files\\BI485 Folder\\Kleinberg_Donovan\\data"
sparrow_raw <- read_tsv(file.path(file_path, "sparrows.txt"), col_names = TRUE)
sparrow_raw
```

Chunk adds rows needed.

```{r}
sparrows <- sparrow_raw %>%
  mutate(box_group = "Sparrows",
         row_order = 1:nrow(.))

View(sparrows)
```

Chunk creates initial boxplot

```{r}
plot_box <- sparrows %>%
  ggplot(aes(x = box_group, y = wingcrd)) +
  geom_boxplot(width = 0.25) +
  labs(x = NULL, y = "Wing Length (mm)")
```

Chunk adds cleveland plot

```{r}
plot_cleveland <- sparrows %>%
  ggplot(aes(x = wingcrd, y = row_order)) +
  geom_point() +
  labs(x = "Wing Length (mm)", y = "Order of the data")
```

Chunk is stitching the plots together

```{r}
plot_box + plot_cleveland + coord_flip()
```
Figure 3

Chunk begins work on figure 3 by creating all plots and assigning them to variables

```{r}
p_culmen <- sparrows %>% 
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = culmen)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

p_culmen

p_nalospi <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = nalospi)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

p_nalospi

p_wt <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wt)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

p_wingcrd <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wingcrd)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

p_tarsus <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = tarsus)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

p_head <- sparrows %>%
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = head)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

p_culmen + p_nalospi + p_wt + p_wingcrd + p_tarsus + p_head + plot_layout(nrow = 2)
```

Answer for question regarding figure 3: It would appear the the head, tarsus and nalospi are the variables with extreme outliers.

Figure 5

Chunk plots out histogram

```{r}
h1 <- sparrows %>%
    filter(Month %in% 6:8) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by = 2))

h1
```

Chunk plots out second facet histograms

```{r}
h2 <- sparrows %>%
  filter(Month %in% 6:8) %>% 
  mutate(Month = case_when(
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August")) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by = 2)) +
  facet_wrap(~ Month, ncol = 1)

h2
```

Chunk puts h1 and h2 together

```{r}
h1 + h2
```

Figure 10

