---
title: "HW 08 Part 1"
author: "Donovan Kleinberg"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


```{r}
library(tidyverse)
library(stringr)
install.packages("ggthemes")
library(ggthemes)
install.packages("RColorBrewer")
install.packages("viridis")
```

Chunk imports data via read_lines

```{r}

beta_gliadin <- read_lines("C:\\Misc. Files\\BI485 Folder\\Kleinberg_Donovan\\data\\beta_gliadin.txt", skip = 1)

beta_gliadin

```

Chunk removes whitespace and digits

```{r}
beta_gliadin_step_one <- beta_gliadin %>%
  str_replace_all("\\s", "\\")

beta_gliadin_step_two <- beta_gliadin_step_one %>%
  str_replace_all("\\d", "\\")

beta_gliadin_step_two
```

Chunk collapses strings

```{r}
beta_gliadin_coll <- beta_gliadin_step_two %>%
  str_c(collapse = "")

beta_gliadin_coll
```

Chunk converts to uppercase

```{r}
beta_gliadin_upper <- beta_gliadin_coll %>%
  str_to_upper()

beta_gliadin_upper

str_length(beta_gliadin_upper)
```

<<<<<<< HEAD
Chunk counts non identified nucleotides

```{r}
str_count(beta_gliadin_upper, c("A", "T", "C", "G"))

### All normal letters add up to 3310, no non-identified nucleotides are in the data set.
```

Chunk counts start and stop codons

```{r}
str_count(beta_gliadin_upper, "ATG")
str_count(beta_gliadin_upper, "TAA|TGA")
```

Chunk begins process to finding promotor

```{r}
caat_box <- "CCAAT"
tata_box <- "CTATAA"

str_count(beta_gliadin_upper, caat_box)
str_count(beta_gliadin_upper, tata_box)

```

Chunk searchs for caat_seq and tata_seq.

```{r}
caat_seq <- "CCAAT.{100,150}ATG"
str_count(beta_gliadin_upper, caat_seq)

tata_seq <- "CTATAA.{50,100}ATG"
str_count(beta_gliadin_upper, tata_seq)
```

Chunk extracts caat sequences

```{r}
caat_seqs_all <- str_extract_all(beta_gliadin_upper, caat_seq, simplify = TRUE)

str_count(caat_seqs_all)
```

Chunk finds and prints promoter region

```{r}
str_which(caat_seqs_all, tata_box)
promoter <- str_subset(caat_seqs_all, tata_box)

promoter
str_count(promoter)
```

Part 2

Ants

```{r}
ants <- read.csv("C:\\Misc. Files\\BI485 Folder\\Kleinberg_Donovan\\data\\ants.csv")

ants
```

Chunk tidies the ant data

```{r}
tidy_ants <- ants %>% 
  select(-c(seas.code)) %>%
  gather(key = species, value = abundance, matches("\\w\\.\\w", ignore.case = FALSE))

tidy_ants
```

Chunk has the hard part

```{r}
ants_post_hard <- tidy_ants %>%
  mutate(species = str_replace(species, pattern = "(^[A-Z]*\\w{3})\\w{3,}\\.(\\w{4})\\w{2,}", replacement = "\\1\\2"))

ants_post_hard
```

Chunk uses mutate to turn 2010/2011 into ordered factors

```{r}
ants_final_mutate <- ants_post_hard %>%
  mutate(year = factor(year, levels = c("2010", "2011"), ordered = TRUE))

ants_final_mutate
```

Chunk groups and summarizes

```{r}
ants_grouped <- ants_final_mutate %>%
  group_by(year, species)

ants_grouped

ants_summ <- ants_grouped %>%
  summarize(Total_Abundance = sum(abundance))

ants_summ
```

Chunk plots ants data

```{r}
ggplot(data = ants_summ, mapping = aes(x = reorder(species, Total_Abundance), y = Total_Abundance, group = species)) +
  coord_flip() +
  geom_line() +
  geom_point(aes(shape = year), size = 2) +
  labs(x = NULL, y = "Total Abundance") +
  theme_minimal()
```

Part 3

Flowers

```{r}
flower_size <- read.csv("C:\\Misc. Files\\BI485 Folder\\Kleinberg_Donovan\\data\\flower_size.csv")

flower_size
```

Chunk filters out NA

```{r}
flower_size_filtered <- flower_size %>%
  filter(!is.na(Flower.number))

flower_size_filtered
```

Chunk selects

```{r}
flower_sel <- flower_size_filtered %>%
  select(Species = Species, Cor_length = Corolla.length..cm., Throat_width = Corolla.width.throat..cm.)

flower_sel
```

Chunk seperates

```{r}
flower_sep <- flower_sel %>%
  separate(Species, into = c("Genus", "Species"))

flower_sep
```

Chunk groups and summarizes

```{r}
flower_grouped <- flower_sep %>%
  group_by(Genus, Species)

flower_grouped

flower_summ <- flower_grouped %>%
  summarise(Mean_length = mean(Cor_length), Max_length = max(Cor_length), Mean_width = mean(Throat_width), Max_width = max(Throat_width))

flower_summ
```

Chunk imports second data set

```{r}
pollinators_raw <- read.csv("C:\\Misc. Files\\BI485 Folder\\Kleinberg_Donovan\\data\\flower_pollinators.csv", na.strings = c("", "NA"))

pollinators_raw

### Had to add na.strings since otherwise the missing data would be blank and not filter out via the filter command.
```

Chunk uses select, filter and seperate.

```{r}
pollinators_sel <- pollinators_raw %>%  
  select(-c(Color, starts_with("Source")))

pollinators_sel

pollinators_fil <- pollinators_sel %>%
  filter(!is.na(Pollinator))

pollinators_fil
  
pollinators <- pollinators_fil %>%
  separate(ï..Species, into = c("Genus", "Species"))

pollinators
```

```{r}
pollinators_sep <- pollinators %>%
  separate(Pollinator, into = "Syndrome", sep = " |, ")

pollinators_sep
```



