---
title: "HW 06 Part 2: Wrangle II"
author: "Donovan Kleinberg"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


```{r}
library(tidyverse)
library(knitr)
install.packages("smatr")
library(smatr)
```

Darters

```{r}
file_path <- "C:\\Misc. Files\\BI485 Folder\\Kleinberg_Donovan\\data"
darters <- read.delim(file.path(file_path, "darter.txt"))

Column_start <- c(1, 4, 15, 16, 17, 18, 21, 25, 29, 33, 36, 39, 40)
Column_stop <- c(3, 14, 15, 16, 17, 20, 24, 28, 32, 35, 38, 40, 41)
names <- c("specimen id", "species", "riffle", "age", "sex", "length", "depth", "velocity", "temperature", "majsub", "minsub", "majtype", "mintype")
tidy_darters <- read_fwf(file.path(file_path, "darter.txt"), fwf_positions(Column_start, Column_stop, col_names = names), skip = 15)

tidy_darters2 <- tidy_darters %>%
  filter(!species %in% c("zonale", "tetrazonum"))
```

```{r}
tidy_darters2
```

Chunk adds total_substrate column.

```{r}
mutate(tidy_darters2, total_substrate = majsub + minsub)
```

All total substrate values are summed to 1.

Chunk has assigmnets of names and change of length to millimeters.

```{r}
tidy_darters2$riffle[tidy_darters2$riffle==1] <- "Riffle_1"
tidy_darters2$riffle[tidy_darters2$riffle==2] <- "Riffle_2"

tidy_darters2$sex[tidy_darters2$sex=="f"] <- "Female"
tidy_darters2$sex[tidy_darters2$sex=="m"] <- "Male"

tidy_darters2 <- mutate(tidy_darters2, length = length * 10)

tidy_darters2
```

Chunk groups and summarizes the required variables and plots them in a table.

```{r}
darter_means <- tidy_darters2 %>% group_by(species) %>%
  summarise_at(.vars = vars(length, depth, velocity), 
               .funs = c(mean = "mean")) 

darter_means
```

Chunk represents darter plot 1.

```{r}
ggplot(data = tidy_darters2) +
  stat_summary(
    mapping = aes(x = species, y = length),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  ) +
  geom_point(data = darter_means, mapping = aes(x = species, y = length_mean), shape = 15) + 
  coord_flip() +
    facet_wrap(~ riffle) +
   ylab("Length (mm)") + xlab("Species")
```

Discussion for plot 1: Overall on riffle 2 the means seem to be much lower than on riffle wh1 for spectabile and caeruleum and about equal for blennioidies with a larger amount of large outliers. 

Chunk represents darter plot 2.

```{r}
ggplot(data = tidy_darters2, mapping = aes(x = species, y = length)) + 
  geom_point(color = "gray65") +
  stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median,
    geom = "point",
    size = 2.2
  ) +
  stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = mean,
    size = 3,
    geom = "point",
    color = "maroon"
  ) +
  coord_flip() +
  facet_wrap(~riffle) +
  xlab("Species") + ylab("Length(mm)")
```

Stress and Cortisone 

```{r}
names <- c("capture_date", "day", "site", "urban_score", "cortF", "cortB", "sex", "smi", "tarsus", "mass", "basal_cort", "stress_cort")
file_path <- "C:\\Misc. Files\\BI485 Folder\\Kleinberg_Donovan\\data"
tidy_stress <- read_csv(file.path(file_path, "feathers.csv"), locale = locale(decimal_mark = ","), skip = 1, col_names = names)
tidy_stress
```

Chunk creates mean tarsus

```{r}
mean_tarsus <- mean(tidy_stress[["tarsus"]])

mean_tarsus
```

Chunk adds log_mass and log_tarsus

```{r}
tidy_stress <- mutate(tidy_stress, log_mass = log(mass), log_tarsus = log(tarsus))

tidy_stress
```

Chunk runs the sma function along with summaries.

```{r}
major_axis <- sma(log_mass ~ log_tarsus, data = tidy_stress)
summary(major_axis)
```

Chunk has coefficent.

```{r}
coef(major_axis)

b <- 1.864889
```

Chunk adds new_smi

```{r}
tidy_stress <- mutate(tidy_stress, new_smi = mass * (mean_tarsus / tarsus) ^ b)
tidy_stress
```

Discussion for new_smi: The new column now extends to 5 decimal places instead of only one decimal place in the old column.

Chunk groups by site.

```{r}
tidy_stress %>%
  group_by(site)
```

Chunk calculates mean and se

```{r}
mean_basal_cort <- mean(tidy_stress$basal_cort, na.rm = TRUE)
mean_stress_cort <- mean(tidy_stress$stress_cort, na.rm = TRUE)
sd_basal_cort <- sd(tidy_stress$basal_cort, na.rm = TRUE)
sd_stress_cort <- sd(tidy_stress$stress_cort, na.rm = TRUE)
se_basal_cort <- sd_basal_cort/(sqrt(43))
se_stress_cort <- sd_stress_cort/(sqrt(43))
```

Plot 1

```{r}
ggplot(data = tidy_stress, mapping = aes(x = new_smi, y = site)) + 
  geom_boxplot()
```

Discussion for plot 1: Yes, there are some very obvious differences between the sites. CEBC has the hights smi at over 25, Larochelle and Villefollet fall behind and Niort has the lowest ranging betewen 22 and 23.

Plot 2

```{r}
min_basal <- mean_basal_cort - se_basal_cort
max_basal <- mean_basal_cort + se_basal_cort

ggplot(data = tidy_stress, mapping = aes(x = site, y = basal_cort)) + 
  geom_point() +
  geom_errorbar(ymin = min_basal, ymax = max_basal)
```

Plot 3

```{r}
min_stress <- mean_stress_cort - se_stress_cort
max_stress <- mean_stress_cort + se_stress_cort

ggplot(data = tidy_stress, mapping = aes(x = site, y = stress_cort)) + 
  geom_point() +
  geom_linerange(ymin = min_stress, ymax = max_stress)
```


