---
title: "HW 06 Part 1: Complete the sections"
author: "Donovan Kleinberg"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.

## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data transformation](https://r4ds.had.co.nz/transform.html), sections 5-7.

You do not answer any of the questions this week. Just read and run all of the examples.

### Section 5.5: Add new variables with `mutate()`

Type and run all of the examples.

```{r}
library(nycflights13)
library(tidyverse)
```


Example 1:

```{r}
flights_sml <- select(flights, year:day, ends_with("delay"), distance, air_time)

mutate(flights_sml, gain = dep_delay - arr_delay, speed = distance / air_time * 60)
```

Description of example 1: Shows the use of the mutate command to add columns gain and speed.

Example 2:

```{r}
mutate(flights_sml, gain = dep_delay - arr_delay, hours = air_time/60, gain_per_hour = gain / hours)
```

Description of example 2: Further use of the mutate command with some of the columns previously added.

Example 3:

```{r}
transmute(flights, gain = dep_delay - arr_delay, hours = air_time / 60, gain_per_hour = gain / hours)
```

Description of example 3: Use of the transmute command to display only columns that were added.

Example 4:

```{r}
transmute(flights, dep_time, hour = dep_time %/% 100, minute = dep_time %% 100)
```

Description of example 4: Use of arithmetic operators to create dep_time.

Example 5:

```{r}
(x <- 1:10)
lag(x)
lead(x)
```

Description of example 5: Use of the lead and lag commands.

Example 6:

```{r}
x
cumsum(x)
cummean(x)
```

Description for example 6: Use of the cumsum and cummean commands.

Example 7:

```{r}
y <- c(1, 2, 2, NA, 3, 4)
min_rank(y)
min_rank(desc(y))
```

Description of example 7: Use of the minrank commands to rank the data.

Example 8:

```{r}
row_number(y)
dense_rank(y)
percent_rank(y)
cume_dist(y)
```

Description of example 8: More niche types of ranking commands.

Pay careful attention to the useful creation functions in section 5.5.1. 

### Section 5.6: Grouped summaries with `summarise()`

Read and run all of the examples. Pay careful attention to the useful summary functions. Unless you really like baseball, skip the baseball part of 5.6.3.

### Section 5.7: Grouped mutates (and filters)

Read and run all of the examples.

