---
title: "HW 06 Part 2: Wrangle II"
author: "Donovan Kleinberg"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


```{r}
library(tidyverse)
library(knitr)
install.packages("smatr")
library(smatr)
```

Darters

```{r}
file_path <- "C:\\Misc. Files\\BI485 Folder\\Kleinberg_Donovan\\data"
darters <- read.delim(file.path(file_path, "darter.txt"))

Column_start <- c(1, 4, 15, 16, 17, 18, 21, 25, 29, 33, 36, 39, 40)
Column_stop <- c(3, 14, 15, 16, 17, 20, 24, 28, 32, 35, 38, 40, 41)
names <- c("specimen id", "species", "riffle", "age", "sex", "length", "depth", "velocity", "temperature", "majsub", "minsub", "majtype", "mintype")
tidy_darters <- read_fwf(file.path(file_path, "darter.txt"), fwf_positions(Column_start, Column_stop, col_names = names), skip = 15)

tidy_darters2 <- tidy_darters %>%
  filter(!species %in% c("zonale", "tetrazonum"))
```

```{r}
tidy_darters2
```

Chunk adds total_substrate column.

```{r}
mutate(tidy_darters2, total_substrate = majsub + minsub)
```

All total substrate values are summed to 1.

Chunk has assigmnets of names and change of length to millimeters.

```{r}
tidy_darters2$riffle[tidy_darters2$riffle==1] <- "Riffle_1"
tidy_darters2$riffle[tidy_darters2$riffle==2] <- "Riffle_2"

tidy_darters2$sex[tidy_darters2$sex=="f"] <- "Female"
tidy_darters2$sex[tidy_darters2$sex=="m"] <- "Male"

tidy_darters2 <- mutate(tidy_darters2, length = length * 10)

tidy_darters2
```

Chunk groups and summarizes the required variables and plots them in a table.

```{r}
darter_means <- tidy_darters2 %>% group_by(species) %>%
  summarise_at(.vars = vars(length, depth, velocity), 
               .funs = c(mean = "mean")) 

darter_means
```

Chunk represents darter plot 1.

```{r}
ggplot(data = tidy_darters2) +
  stat_summary(
    mapping = aes(x = species, y = length),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  ) +
  geom_point(data = darter_means, mapping = aes(x = species, y = length_mean), shape = 15) + 
  coord_flip() +
    facet_wrap(~ riffle) +
   ylab("Length (mm)") + xlab("Species")
```

Discussion for plot 1: Overall on riffle 2 the means seem to be much lower than on riffle wh1 for spectabile and caeruleum and about equal for blennioidies with a larger amount of large outliers. 

Chunk represents darter plot 2.

```{r}
ggplot(data = tidy_darters2, mapping = aes(x = species, y = length)) + 
  geom_point(color = "gray65") +
  stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median,
    geom = "point",
    size = 2.2
  ) +
  stat_summary(
    fun.ymin = min,
    fun.ymax = max,
    fun.y = mean,
    size = 3,
    geom = "point",
    color = "maroon"
  ) +
  coord_flip() +
  facet_wrap(~riffle) +
  xlab("Species") + ylab("Length(mm)")
```

Stress and Cortisone 



