---
title: "HW 06 Part 1: Complete the sections"
author: "Donovan Kleinberg"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- Change "your name" in the YAML header above to your name.

- As usual, enter the examples in code chunks and run them, unless told otherwise.

## Chapter 5: Data transformation

Read [R4ds Chapter 5: Data transformation](https://r4ds.had.co.nz/transform.html), sections 5-7.

You do not answer any of the questions this week. Just read and run all of the examples.

### Section 5.5: Add new variables with `mutate()`

Type and run all of the examples.

```{r}
library(nycflights13)
library(tidyverse)
```


Example 1:

```{r}
flights_sml <- select(flights, year:day, ends_with("delay"), distance, air_time)

mutate(flights_sml, gain = dep_delay - arr_delay, speed = distance / air_time * 60)
```

Description of example 1: Shows the use of the mutate command to add columns gain and speed.

Example 2:

```{r}
mutate(flights_sml, gain = dep_delay - arr_delay, hours = air_time/60, gain_per_hour = gain / hours)
```

Description of example 2: Further use of the mutate command with some of the columns previously added.

Example 3:

```{r}
transmute(flights, gain = dep_delay - arr_delay, hours = air_time / 60, gain_per_hour = gain / hours)
```

Description of example 3: Use of the transmute command to display only columns that were added.

Example 4:

```{r}
transmute(flights, dep_time, hour = dep_time %/% 100, minute = dep_time %% 100)
```

Description of example 4: Use of arithmetic operators to create dep_time.

Example 5:

```{r}
(x <- 1:10)
lag(x)
lead(x)
```

Description of example 5: Use of the lead and lag commands.

Example 6:

```{r}
x
cumsum(x)
cummean(x)
```

Description for example 6: Use of the cumsum and cummean commands.

Example 7:

```{r}
y <- c(1, 2, 2, NA, 3, 4)
min_rank(y)
min_rank(desc(y))
```

Description of example 7: Use of the minrank commands to rank the data.

Example 8:

```{r}
row_number(y)
dense_rank(y)
percent_rank(y)
cume_dist(y)
```

Description of example 8: More niche types of ranking commands.

Pay careful attention to the useful creation functions in section 5.5.1. 

### Section 5.6: Grouped summaries with `summarise()`

Read and run all of the examples. Pay careful attention to the useful summary functions. Unless you really like baseball, skip the baseball part of 5.6.3.

Example 1:


```{r}
by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE))

delay <- filter(delay, count > 20, dest != "HNL")

ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```

Description for example 1: Creating a graph to explore the relationship between dist and delay while filtering out noise.

Example 2:


```{r}
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(count = n(), dist = mean(distance, na.rm = TRUE), delay = mean(arr_delay, na.rm = TRUE)) %>% 
  filter(count > 20, dest != "HNL")
```

Description for example 2: Same as before but using the almighty pipe.

Example 3:

```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```

Description for example 3: Lack of the na.rm argument causes all the data with missing entries to be displayed.

Example 4:

```{r}
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay, na.rm = TRUE))
```

Description for example 4: Use of na.rm argument removes the missing data points and observations with them.

Example 5:

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```

Desc. for 5: Creation of some variables for later.

Example 6:

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(delay = mean(arr_delay))

ggplot(data = delays, mapping = aes(x = delay)) + 
  geom_freqpoly(binwidth = 10)
```

Desc. for 6: Graph  of planes grouped by tail number against their delay times.

Example 7:

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(delay = mean(arr_delay, na.rm = TRUE), n = n())

ggplot(data = delays, mapping = aes(x = n, y = delay)) + 
  geom_point(alpha = 1/10)
```

Desc. for 7: Sorting against the mean shows that only a few planes have extremely high delay times.

Example 8:

```{r}
delays %>% 
  filter(n > 25) %>% 
  ggplot(mapping = aes(x = n, y = delay)) + 
    geom_point(alpha = 1/10)
```

Desc. for 8: Filtering out observations with few numbers to get rid of small influences on the graph.

Example 9:

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    avg_delay1 = mean(arr_delay),
    avg_delay2 = mean(arr_delay[arr_delay > 0])
  )
```

Desc. for 9: Calculating some averages of negative and positive delays.

Example 10:

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(first = min(dep_time), last = max(dep_time))
```

Desc. for 10: Finds out what times the first and last flights left.

Example 11:

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  mutate(r = min_rank(desc(dep_time))) %>% 
  filter(r %in% range(r))
```

Desc. for 11: Grouped by data, ranks on dep time.

Example 12:

```{r}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(carriers = n_distinct(carrier)) %>% 
  arrange(desc(carriers))
```

Desc. for 12: Using n_distinct to count the number of unique carriers at an airport.

Example 13:

```{r}
not_cancelled %>% 
  count(dest)
```

Desc for 13: Just a simple count.

Example 14:

```{r}
not_cancelled %>% 
  count(tailnum, wt = distance)
```

Desc. for 14: Usage of a wt to sum how far each plane has flown.

Example 15:

```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(n_early = sum(dep_time < 500))

not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(hour_perc = mean(arr_delay > 60))
```

Desc. for 15: Uses < and > sums to determine which flights left before 5am and which were delayed by more than an hour.

Example 16:

```{r}
daily <- group_by(flights, year, month, day)
(per_day   <- summarise(daily, flights = n()))
(per_month <- summarise(per_day, flights = sum(flights)))
(per_year  <- summarise(per_month, flights = sum(flights)))
```

Desc. for 16: Using multiple summaries to "peel" layers off a dataset.

Example 17:

```{r}
daily %>% 
  ungroup() %>%             
  summarise(flights = n())
```

Desc. for 17: Ungroups the variable.

### Section 5.7: Grouped mutates (and filters)

Read and run all of the examples.

